\textbf{Beasts of Prey or Rational Animals?\\ Private Governance in Brazil's \emph{Jogo do Bicho}}

### 2018-09-20 paper-bicho/bicho.tex
update intro3

Yet, the \emph{jogo do bicho} has survived. The animal game has outlasted more than 30 Brazilian presidents and thrived under military regimes and democratic governments alike \citep{jupiara2015poroes}. But more than a act of defiance, the \emph{jogo do bicho} is a successful capitalist enterprise \citep{labronici2014sorteio, magalhaes2005ganhou}. A recent study by Fundação Getúlio Vargas, a Brazilian think tank, affirmed that the \emph{jogo do bicho} earns from BRL 1.3 to BRL 2.8 billion per year (USD 400 to USD 850 million), making it the largest clandestine gambling game in the world.\footnote{See \href{https://www.huffpostbrasil.com/2015/12/07/legalizacao-jogos-brasil_n_8737212.html}{\texttt{https://www.huffpostbrasil.com/2015/12/07/legalizacao-jogos-brasil\_n\_8737212.html}} (in Portuguese). Access: August 2018.} \citet[171]{schneider1996brazil} estimated that in the 1990s, the game furnished about 50,000 jobs in the Rio de Janeiro city alone, almost the same number of employees that the oil giant Petrobras had in 2011 \citep{exame2013petrobras}.\footnote{In 1966, Time Magazine wrote that the \emph{jogo do bicho} was `the largest single industry in Latin America' and employed about 1\% of the Brazilian workforce. See \href{http://content.time.com/time/magazine/article/0,9171,842527-1,00.html}{\texttt{http://content.time.com/time/magazine/article/0,9171,842527-1,00.html}}. Access: August 2018.}

In this article I offer a rational choice interpretation of the \emph{jogo do bicho}. More specifically, I discuss how \emph{bicheiros} adopt rational strategies to solve collective action problems and maximise their political leverage. I use the case of Rio de Janeiro to illustrate the lottery's profit abilities and its effect in political corruption. The lottery financiers have long been sponsors of Rio de Janeiro's world-famous Carnival Parade, and several of the country's most high-profile \emph{bicheiros} hail from that state. Furthermore, in no other part of Brazil the patron-client relationship between \emph{bicheiros} and politicians has been so widely publicised, with game financiers openly funding cultural activities to co-opt local politicians attracted by the electoral gains offered by the festival \citep{cavalcanti2006carnaval, queiroz1992carnaval}.

>>>>

Yet the \emph{jogo do bicho} has survived. The animal game has outlasted more than 30 Brazilian presidents and thrived under military regimes and democratic governments alike \citep{jupiara2015poroes}. But more than a act of defiance, the \emph{jogo do bicho} is a successful capitalist enterprise \citep{labronici2014sorteio, magalhaes2005ganhou}. A recent study by Fundação Getúlio Vargas, a Brazilian think tank, affirmed that the \emph{jogo do bicho} earns from BRL 1.3 to BRL 2.8 billion per year (USD 400 to USD 850 million), making it the largest clandestine gambling game in the world.\footnote{See \href{https://www.huffpostbrasil.com/2015/12/07/legalizacao-jogos-brasil_n_8737212.html}{\texttt{https://www.huffpostbrasil.com/2015/12/07/legalizacao-jogos-brasil\_n\_8737212.html}} (in Portuguese). Access: August 2018.} \citet[171]{schneider1996brazil} estimated that in the 1990s, the game furnished about 50,000 jobs in the Rio de Janeiro city alone, almost the same number of employees that the oil giant Petrobras had in 2011 \citep{exame2013petrobras}.\footnote{In 1966, Time Magazine wrote that the \emph{jogo do bicho} was `the largest single industry in Latin America' and employed about 1\% of the Brazilian workforce. See \href{http://content.time.com/time/magazine/article/0,9171,842527-1,00.html}{\texttt{http://content.time.com/time/magazine/article/0,9171,842527-1,00.html}}. Access: August 2018.}

In this article I offer a rational choice interpretation of the \emph{jogo do bicho}. More specifically, I discuss how \emph{bicheiros} adopt rational strategies to solve collective action problems and maximise their political leverage. I use the case of Rio de Janeiro to illustrate the lottery's profit abilities and its effect in political corruption. The lottery financiers have long been major sponsors of Rio de Janeiro's world-famous Carnival Parade, and several of the country's most high-profile \emph{jogo do bicho} entrepreneurs hail from that state. Furthermore, in no other part of Brazil the patron-client relationship between \emph{bicheiros} and politicians has been so widely publicised.

 to co-opt local politicians attracted by the electoral gains offered by the festival \citep{cavalcanti2006carnaval, queiroz1992carnaval}.
